---
title: "A/B Test Analysis: Checkout Flow Redesign"
author: "Growth Analytics"
date: today
format:
  html:
    toc: true
    code-fold: true
    theme: cosmo
  pdf:
    documentclass: article
execute:
  echo: true
  warning: false
  cache: true
---

## Background

This report analyses a 28-day A/B test evaluating whether a simplified
3-step checkout flow (treatment) outperforms the existing 5-step flow (control)
on conversion rate and average order value.

**Hypothesis:** Reducing friction in checkout increases conversion without
reducing average order value.

## Setup

```{python}
import pandas as pd
import numpy as np
from scipy import stats
from scipy.stats import chi2_contingency, mannwhitneyu
import matplotlib.pyplot as plt
import matplotlib.patches as mpatches
from statsmodels.stats.proportion import proportions_ztest
from statsmodels.stats.power import zt_ind_solve_power
import warnings
warnings.filterwarnings('ignore')

np.random.seed(2024)

ALPHA          = 0.05
MDE            = 0.02    # minimum detectable effect (2 pp)
TEST_DURATION  = 28      # days
TRAFFIC_SPLIT  = 0.50
```

```{r setup-r}
#| message: false
library(tidyverse)
library(scales)

# Load the same data for R-side visualisations
ab_data <- read_csv("ab_test_results.csv", show_col_types = FALSE)
```

## Data Generation (Simulation)

```{python}
n_control   = 24_800
n_treatment = 25_200

control_cr    = 0.042
treatment_cr  = 0.056
control_aov   = 87.50
treatment_aov = 89.20

control_converted   = int(n_control   * control_cr)
treatment_converted = int(n_treatment * treatment_cr)

control_orders = pd.DataFrame({
    'group': 'control',
    'converted': ([1] * control_converted + [0] * (n_control - control_converted)),
    'order_value': (list(np.random.gamma(3, control_aov / 3, control_converted)) +
                    [0] * (n_control - control_converted))
})

treatment_orders = pd.DataFrame({
    'group': 'treatment',
    'converted': ([1] * treatment_converted + [0] * (n_treatment - treatment_converted)),
    'order_value': (list(np.random.gamma(3, treatment_aov / 3, treatment_converted)) +
                    [0] * (n_treatment - treatment_converted))
})

ab = pd.concat([control_orders, treatment_orders], ignore_index=True).sample(frac=1, random_state=42)
print(f"Total visitors: {len(ab):,}")
print(ab.groupby('group')['converted'].agg(['sum', 'mean']).rename(
    columns={'sum': 'conversions', 'mean': 'conversion_rate'}))
```

## Statistical Tests

### Conversion Rate

```{python}
ctrl  = ab[ab['group'] == 'control']
trt   = ab[ab['group'] == 'treatment']

counts   = np.array([trt['converted'].sum(),  ctrl['converted'].sum()])
nobs     = np.array([len(trt), len(ctrl)])
z_stat, p_value = proportions_ztest(counts, nobs, alternative='larger')

relative_lift = (trt['converted'].mean() - ctrl['converted'].mean()) / ctrl['converted'].mean()

print(f"Control CR:    {ctrl['converted'].mean():.2%}")
print(f"Treatment CR:  {trt['converted'].mean():.2%}")
print(f"Relative lift: {relative_lift:+.1%}")
print(f"Z-statistic:   {z_stat:.3f}")
print(f"P-value:       {p_value:.4f}")
print(f"Significant:   {p_value < ALPHA}")
```

### Average Order Value (Among Converters)

```{python}
ctrl_aov = ctrl[ctrl['converted'] == 1]['order_value']
trt_aov  = trt[trt['converted']  == 1]['order_value']

u_stat, p_mann = mannwhitneyu(trt_aov, ctrl_aov, alternative='two-sided')

print(f"Control AOV:   ${ctrl_aov.mean():.2f}")
print(f"Treatment AOV: ${trt_aov.mean():.2f}")
print(f"Mann-Whitney U: {u_stat:.0f},  p = {p_mann:.4f}")
print(f"AOV difference significant: {p_mann < ALPHA}")
```

### Revenue Impact

```{python}
daily_visitors    = (n_control + n_treatment) / TEST_DURATION
ctrl_daily_rev    = daily_visitors * TRAFFIC_SPLIT * ctrl['converted'].mean() * ctrl_aov.mean()
trt_daily_rev     = daily_visitors * TRAFFIC_SPLIT * trt['converted'].mean() * trt_aov.mean()
annual_uplift_est = (trt_daily_rev - ctrl_daily_rev) * 365

print(f"Estimated annual revenue uplift: ${annual_uplift_est:,.0f}")
```

## Segmentation

```{python}
device_type = np.random.choice(['mobile', 'desktop', 'tablet'],
                               len(ab), p=[0.58, 0.35, 0.07])
ab['device'] = device_type

seg = ab.groupby(['group', 'device'])['converted'].agg(['sum', 'count', 'mean'])
seg.columns = ['conversions', 'visitors', 'conversion_rate']
print(seg.round(4))
```

## Visualisations

```{r plots}
#| fig-width: 10
#| fig-height: 5
ggplot(ab_data, aes(x = group, y = converted, fill = group)) +
  stat_summary(fun = mean, geom = "bar", alpha = 0.85) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2) +
  scale_fill_manual(values = c("control" = "#6c757d", "treatment" = "#0d6efd")) +
  scale_y_continuous(labels = percent_format()) +
  labs(title = "Conversion Rate by Group",
       subtitle = "Error bars show ±1 SE",
       x = NULL, y = "Conversion Rate") +
  theme_minimal(base_size = 13) +
  theme(legend.position = "none")
```

## Conclusion

```{python}
print("=" * 50)
print("TEST CONCLUSION")
print("=" * 50)
print(f"Result:         {'SHIP' if p_value < ALPHA else 'DO NOT SHIP'}")
print(f"CR lift:        {relative_lift:+.1%}  (p={p_value:.4f})")
print(f"AOV impact:     {'Neutral' if p_mann >= ALPHA else 'Significant'}")
print(f"Annual uplift:  ${annual_uplift_est:,.0f}")
print("Recommendation: Roll out simplified checkout to 100% of users.")
```
